---
title: "torchvision classes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{torchvision classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

What a torchvision dataset hands back is a plain R list with an S3 class glued
onto it, and that class is what the rest of the package dispatches on. It decides
which transform will accept the object, what the object holds once that transform
has run, and which drawing helper can render it. This article goes through those
classes and shows what is inside each of them.

The code is not evaluated when the website is built, since most of it needs a
dataset on disk.

```{r setup}
library(torchvision)
library(torch)
```

## Why the classes exist

An item is a pair that has to stay consistent. Rotate the image and the boxes
have to turn with it; crop it and the masks have to be cropped the same way. The
class on the item is what records which kind of annotation it carries, so that
one function name can do the right thing for each of them and refuse the ones it
cannot handle, rather than return a mismatched pair.

The target transforms used to read the fields of `y` to work out what they had
been handed. Dispatching on a class instead makes the contract explicit: a
target that has lost its class fails at the call, rather than producing a wrong
result two steps later. It is also what lets the drawing helpers take an item
whole, since `draw_bounding_boxes(item)` knows where the image, the boxes and
the labels are only because the class says so.

## Datasets, items and targets

Indexing a dataset returns an *item*: a list with `x`, the image, and `y`,
everything the dataset knows about that image. `y` is called the *target*.

![A dataset is a deck of items; each item holds an image in `$x` and its annotations in `$y`.](assets/dataset-item.svg)

A dataset takes up to three transform arguments, and they reach different depths
of that structure. `transform` is handed `x` alone, `target_transform` is handed
`y` alone, and `item_transform` is handed the item as a whole, once the other two
have run. Classification datasets only offer the first two. Detection and
segmentation datasets offer all three, because an image cannot be rotated or
cropped without moving its boxes and masks with it.

```{r}
ds <- coco_detection_dataset(
  train = FALSE,
  download = TRUE,
  transform = transform_to_tensor,
  item_transform = function(item) item_transform_resize(item, c(480, 640))
)
```

`transform_to_tensor` is doing more work there than it looks. A dataset returns
`x` in whatever form it was read in, which is the `(H, W, 3)` channel-last array
of `base_loader()` for most of them and a magick image for
`cityscapes_dataset()`. The item transforms and the drawing helpers all want the
`(C, H, W)` tensor that torch models take, so a detection or segmentation
dataset is nearly always built with `transform = transform_to_tensor`.

The dataset itself is classed by task, with `object_detection_dataset` or
`segmentation_dataset` inserted ahead of `dataset` so that task methods win over
the generic ones while the dataset keeps its own name first. That class is what
lets you hand a whole dataset to an item transform:

```{r}
ds <- item_transform_random_horizontal_flip(ds, p = 0.5)
```

which composes the flip into the dataset's `item_transform` rather than applying
it once.

## Classification and caption items

The simplest items carry no class at all. `mnist_dataset()`, `cifar10_dataset()`,
`eurosat_dataset()` and the other classification datasets return a bare list
whose `y` is the class index:

```{r}
mnist <- mnist_dataset(root = tempdir(), download = TRUE, transform = transform_to_tensor)
str(mnist[1])
#> List of 2
#>  $ x:Float [1:1, 1:28, 1:28]
#>  $ y: int 6
```

`flickr8k_caption_dataset()` and `flickr30k_caption_dataset()` have the same
shape, with a character vector of captions in `y`. Nothing dispatches on these
items: an image transform goes in through `transform`, and there is no geometry
in the target to keep in sync with it.

## Detection items

`coco_detection_dataset()`, `pascal_detection_dataset()` and the
`rf100_*_collection()`s give their items the class `image_with_bounding_box` and
their target the class `object_detection_target`.

![An `image_with_bounding_box` item and the fields of its `object_detection_target`.](assets/detection-item.svg)

| field | contents |
|---|---|
| `$boxes` | `(N, 4)` tensor, one row per object, as `(x_min, y_min, x_max, y_max)` in pixels |
| `$labels` | the class name of each box |
| `$area` | `(N)` float tensor, the area of each annotation |
| `$iscrowd` | `(N)` boolean tensor, `TRUE` where the annotation covers a group of objects rather than one |
| `$image_height`, `$image_width` | the size `x` had when the boxes were read |

Only `$boxes` is guaranteed, since that is the field the class is validated on.
The rest is whatever the annotations happened to carry: the table above is a
COCO target, while an RF100 one holds `$boxes`, `$labels` and an `$image_id` and
nothing else.

`draw_bounding_boxes()` has a method for the item, so it finds `$x`,
`$y$boxes` and `$y$labels` on its own:

```{r}
boxed <- draw_bounding_boxes(ds[1])
tensor_image_browse(boxed)
```

## Rotated box items

Four numbers per box cannot describe a box that has been turned, so
`item_transform_rotate()` and `item_transform_affine()` return an
`image_with_rotated_box` instead of the item they were given:

```{r}
class(item_transform_rotate(ds[1], angle = 15))
#> [1] "image_with_rotated_box" "list"
```

![An `image_with_rotated_box` item: same target fields, but `$boxes` has a fifth column.](assets/rotated-box-item.svg)

The target keeps its `object_detection_target` class and its fields, but
`$boxes` gains a fifth column holding the accumulated angle in degrees, the
xyxyr format that `box_xyxy_to_xyxyr()` and `target_transform_rotate()` produce.
`draw_bounding_boxes()` has a second method for this class which draws each box
as a polygon through magick rather than as an upright rectangle.

Every item transform has a method for rotated-box items, except
`item_transform_perspective()`.

## Segmentation items

`coco_segmentation_dataset()`, `pascal_segmentation_dataset()`,
`cityscapes_dataset()`, `oxfordiiitpet_segmentation_dataset()` and
`rf100_peixos_segmentation_dataset()` return `image_with_segmentation_mask`
items whose target is a `segmentation_target`.

![An `image_with_segmentation_mask` item and its `segmentation_target`.](assets/segmentation-item.svg)

No field is required here. Segmentation annotations differ far more between
datasets than detection ones do, so each dataset passes on what it reads:
polygons in `$segmentation` for COCO, `$polygon`, `$color`, `$semantic` and
`$instance` for Cityscapes depending on the requested `target_type`, one-hot
`$masks` for Oxford-IIIT Pet and Pascal VOC. `$labels` is not uniform either:
the detection datasets name their classes, while
`pascal_segmentation_dataset()` numbers the planes of `$masks` that are not
empty.

What `draw_segmentation_masks()` and the segmentation models want is `$masks`,
an `(N, H, W)` tensor with one plane per annotated class. Two target transforms
build it from raw annotations:

| transform | reads |
|---|---|
| `target_transform_coco_masks()` | `$segmentation`, the COCO polygons |
| `target_transform_trimap_masks()` | `$trimap`, the three-value Oxford-IIIT Pet mask |

```{r}
seg_ds <- coco_segmentation_dataset(
  train = FALSE,
  download = TRUE,
  transform = transform_to_tensor,
  target_transform = target_transform_coco_masks
)
masked <- draw_segmentation_masks(seg_ds[1])
```

`coco_segmentation_dataset()` only promotes its items to
`image_with_segmentation_mask` once `$masks` is there, so without that target
transform they stay `image_with_bounding_box`. `cityscapes_dataset()` has no
target transform of its own, and its targets have to be turned into `$masks` by
hand before they can be drawn.

## Slicing an image: `sahi_split`

SAHI, Slicing Aided Hyper Inference, cuts a large image into overlapping tiles,
runs the model on each and puts the results back together. The tiling itself is
an object: `prepare_sahi_split()` measures an image and returns a `sahi_split`.

![A `sahi_split`, drawn as one card rather than a deck, because it is computed once and reused across images.](assets/sahi-split.svg)

```{r}
split <- prepare_sahi_split(c(1000, 1200), size = c(512, 512))
length(split$crop_windows)
#> [1] 9
```

It holds `$crop_windows`, one `top`/`left`/`height`/`width` window per tile, and
the `$size`, `$overlap_size_ratio`, `$image_height` and `$image_width` it was
computed from. Keeping it as its own object is what lets the same tiling be
applied to an image and to its target, so the two stay aligned:

```{r}
tiled <- ds[1]
split <- prepare_sahi_split(tiled$x, size = c(256, 256))

crops <- transform_sahi_crop(tiled$x, split)
targets <- target_transform_sahi_crop(tiled$y, split)
```

`transform_sahi_crop()` returns the tiles stacked as an `(N, C, H, W)` batch.
`target_transform_sahi_crop()` returns one target per tile, each box clipped to
the tile and shifted into its coordinates, and `$image_height` and
`$image_width` set to the size of the tile rather than of the original image. A
box is kept only if what survives the clip is at least `min_area_ratio` of its
original area, a tenth by default, so a box grazing the edge of a tile is
dropped from that tile rather than kept as a sliver.

Unlike the item transforms, this pair turns one item into many, so there is no
`item_transform_sahi_crop()` doing both halves at once.

## Which function takes which class

| family | dispatches on | acts on |
|---|---|---|
| `transform_*()` | `torch_tensor` | the image alone |
| `prepare_sahi_split()` | `numeric`, `torch_tensor`, `magick-image`, `dataset` | measures an image, returns a `sahi_split` |
| `target_transform_resize()`, `target_transform_rotate()`, `target_transform_affine()`, `target_transform_sahi_crop()` | `object_detection_target` | the target alone |
| `target_transform_coco_masks()`, `target_transform_trimap_masks()` | `segmentation_target` | the target alone |
| `item_transform_*()` | `image_with_bounding_box`, `image_with_rotated_box`, `image_with_segmentation_mask`, `dataset` | image and target together |
| `draw_bounding_boxes()` | `torch_tensor`, `image_with_bounding_box`, `image_with_rotated_box` | draws boxes on the image |
| `draw_segmentation_masks()` | `torch_tensor`, `image_with_segmentation_mask` | overlays masks on the image |

The two groups of target transforms are disjoint, which is why they are two rows
and not one. The first group moves boxes and is written for detection targets
only; the second builds `$masks` out of the annotations a segmentation dataset
carries. Crossing them is an error rather than a silent no-op, so
`target_transform_rotate()` on a `segmentation_target`, or
`target_transform_coco_masks()` on an `object_detection_target`, stops with
`not implemented for` and the class it was handed.

Both groups also accept bare lists with the fields their transforms require.
Transforming a bare list preserves its list class.

The image transforms are written for `torch_tensor`. Five of them also take a
magick image, `transform_crop()`, `transform_hflip()`, `transform_resize()`,
`transform_sahi_crop()` and `transform_to_tensor()`, and `transform_to_tensor()`
alone takes an `array`, a `matrix` or a `list`, which is how an image coming off
a loader gets into a tensor in the first place.

The item transforms are the most uniform family: each of them takes all three
item classes and a `dataset`, apart from `item_transform_perspective()`, which
has no rotated-box method. Among the target transforms only
`target_transform_rotate()` takes a dataset as well. The reference index lists
each of these families in full.

`draw_keypoints()` is not generic: it takes an image tensor and an `(N, K, 2)`
keypoint tensor.

Calling any of these on something they have no method for stops with an error
naming the class it was given, which is usually a sign that a target lost its
class somewhere along a pipeline.

## Building an item by hand

Target transforms accept a bare list of annotations. When building a complete
item for item transforms or drawing functions, give the item its task class;
the example also gives the target its corresponding class:

```{r}
img <- torch_randint(0, 255, size = c(3, 120, 160))$to(dtype = torch_uint8())

y <- list(
  boxes = torch_tensor(rbind(c(10, 20, 60, 90), c(70, 30, 150, 110)), dtype = torch_float()),
  labels = c("cat", "dog"),
  image_height = 120,
  image_width = 160
)
class(y) <- c("object_detection_target", "list")

item <- list(x = img, y = y)
class(item) <- c("image_with_bounding_box", "list")

str(item)
#> List of 2
#>  $ x:Byte [1:3, 1:120, 1:160]
#>  $ y:List of 4
#>   ..$ boxes       :Float [1:2, 1:4]
#>   ..$ labels      : chr [1:2] "cat" "dog"
#>   ..$ image_height: num 120
#>   ..$ image_width : num 160
#>   ..- attr(*, "class")= chr [1:2] "object_detection_target" "list"
#>  - attr(*, "class")= chr [1:2] "image_with_bounding_box" "list"
```

A segmentation target is built the same way, with `segmentation_target` and
`image_with_segmentation_mask`.

## What models return

Model outputs are plain lists and carry no class of their own. The object
detection models return `$detections`, one element per image of the batch, each
with `$boxes`, `$labels` and `$scores`; the semantic segmentation models return
`$out`, the per-class logits. The
[Visualization utilities](examples/visualization-utilities.html) article covers
turning either of them into something the drawing helpers can render.

## Comparison with torchvision TVTensors (advanced)

If you have used the v2 transforms of the Python torchvision, these classes
cover the same ground as
[TVTensors](https://docs.pytorch.org/vision/main/auto_examples/transforms/plot_transforms_getting_started.html#videos-boxes-masks-keypoints),
but they are attached one level further out.

A TVTensor is a subclass of `torch.Tensor`. `Image`, `Video`, `BoundingBoxes`,
`Mask` and `KeyPoints` are tensors that know what they are, and a
`BoundingBoxes` carries its `format` and its `canvas_size` as attributes of the
tensor itself. A v2 transform walks an arbitrary nested sample, a dict, a list
or a tuple, dispatches on the type of each element it finds, hands back the same
structure, and passes anything it does not recognise through untouched.

Here the class goes on the list instead. `$x` and `$y$boxes` are ordinary
`torch_tensor`s, and what carries the meaning is `image_with_bounding_box` on
the item and `object_detection_target` on `y`. Four differences follow from
that:

- An item is always `list(x, y)` rather than any nesting you like, and the three
  transform families say which half of it they act on instead of one transform
  traversing the whole sample.
- Box format is not an attribute but the width of `$boxes`: four columns are
  xyxy, five are xyxyr. What `canvas_size` holds lives in `$image_height` and
  `$image_width` on the target.
- A change of geometry changes the class of the item, so
  `item_transform_rotate()` hands back an `image_with_rotated_box`. On the
  Python side the object stays a `BoundingBoxes` and the distinction lives in
  its `format`.
- There is no heuristic for unwrapped input. A v2 transform whose sample holds
  no `Image` will treat the first plain tensor it finds as the image; here a
  list without its class is an error, not a guess.

Two of the five TVTensors have no counterpart here. There is no video support at
all, and keypoints are not a class: `draw_keypoints()` takes an image tensor and
a keypoint tensor and nothing dispatches on either.
