---
title: "torchvision classes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{torchvision classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

What a torchvision dataset hands back is a plain R list with an S3 class glued
onto it, and that class is what the rest of the package dispatches on. It decides
which transform will accept the object, what the object holds once that transform
has run, and which drawing helper can render it. This article goes through those
classes and shows what is inside each of them.

The code is not evaluated when the website is built, since most of it needs a
dataset on disk.

```{r setup}
library(torchvision)
library(torch)
```

## Datasets, items and targets

Indexing a dataset returns an *item*: a list with `x`, the image, and `y`,
everything the dataset knows about that image. `y` is called the *target*.

![A dataset is a deck of items; each item holds an image in `$x` and its annotations in `$y`.](assets/dataset-item.svg)

A dataset takes up to three transform arguments, and they reach different depths
of that structure. `transform` is handed `x` alone, `target_transform` is handed
`y` alone, and `item_transform` is handed the item as a whole, once the other two
have run. Classification datasets only offer the first two. Detection and
segmentation datasets offer all three, because an image cannot be rotated or
cropped without moving its boxes and masks with it.

```{r}
ds <- coco_detection_dataset(
  train = FALSE,
  download = TRUE,
  transform = transform_to_tensor,
  item_transform = function(item) item_transform_resize(item, c(480, 640))
)
```

`transform_to_tensor` is doing more work there than it looks. A dataset returns
`x` in whatever form it was read in, which is the `(H, W, 3)` channel-last array
of `base_loader()` for most of them and a magick image for
`cityscapes_dataset()`. The item transforms and the drawing helpers all want the
`(C, H, W)` tensor that torch models take, so a detection or segmentation
dataset is nearly always built with `transform = transform_to_tensor`.

The dataset itself is classed by task, with `object_detection_dataset` or
`segmentation_dataset` inserted ahead of `dataset` so that task methods win over
the generic ones while the dataset keeps its own name first. That class is what
lets you hand a whole dataset to an item transform:

```{r}
ds <- item_transform_random_horizontal_flip(ds, p = 0.5)
```

which composes the flip into the dataset's `item_transform` rather than applying
it once.

## Classification and caption items

The simplest items carry no class at all. `mnist_dataset()`, `cifar10_dataset()`,
`eurosat_dataset()` and the other classification datasets return a bare list
whose `y` is the class index:

```{r}
mnist <- mnist_dataset(root = tempdir(), download = TRUE, transform = transform_to_tensor)
str(mnist[1])
#> List of 2
#>  $ x:Float [1:1, 1:28, 1:28]
#>  $ y: int 6
```

`flickr8k_caption_dataset()` and `flickr30k_caption_dataset()` have the same
shape, with a character vector of captions in `y`. Nothing dispatches on these
items: an image transform goes in through `transform`, and there is no geometry
in the target to keep in sync with it.

## Detection items

`coco_detection_dataset()`, `pascal_detection_dataset()` and the
`rf100_*_collection()`s give their items the class `image_with_bounding_box` and
their target the class `object_detection_target`.

![An `image_with_bounding_box` item and the fields of its `object_detection_target`.](assets/detection-item.svg)

| field | contents |
|---|---|
| `$boxes` | `(N, 4)` tensor, one row per object, as `(x_min, y_min, x_max, y_max)` in pixels |
| `$labels` | the class name of each box |
| `$area` | `(N)` float tensor, the area of each annotation |
| `$iscrowd` | `(N)` boolean tensor, `TRUE` where the annotation covers a group of objects rather than one |
| `$image_height`, `$image_width` | the size `x` had when the boxes were read |

Only `$boxes` is guaranteed, since that is the field the class is validated on.
The rest is whatever the annotations happened to carry: the table above is a
COCO target, while an RF100 one holds `$boxes`, `$labels` and an `$image_id` and
nothing else.

`draw_bounding_boxes()` has a method for the item, so it finds `$x`,
`$y$boxes` and `$y$labels` on its own:

```{r}
boxed <- draw_bounding_boxes(ds[1])
tensor_image_browse(boxed)
```

## Rotated box items

Four numbers per box cannot describe a box that has been turned, so
`item_transform_rotate()` and `item_transform_affine()` return an
`image_with_rotated_box` instead of the item they were given:

```{r}
class(item_transform_rotate(ds[1], angle = 15))
#> [1] "image_with_rotated_box" "list"
```

![An `image_with_rotated_box` item: same target fields, but `$boxes` has a fifth column.](assets/rotated-box-item.svg)

The target keeps its `object_detection_target` class and its fields, but
`$boxes` gains a fifth column holding the accumulated angle in degrees, the
xyxyr format that `box_xyxy_to_xyxyr()` and `target_transform_rotate()` produce.
`draw_bounding_boxes()` has a second method for this class which draws each box
as a polygon through magick rather than as an upright rectangle.

Every item transform has a method for rotated-box items, except
`item_transform_perspective()`.

## Segmentation items

`coco_segmentation_dataset()`, `pascal_segmentation_dataset()`,
`cityscapes_dataset()`, `oxfordiiitpet_segmentation_dataset()` and
`rf100_peixos_segmentation_dataset()` return `image_with_segmentation_mask`
items whose target is a `segmentation_target`.

![An `image_with_segmentation_mask` item and its `segmentation_target`.](assets/segmentation-item.svg)

No field is required here. Segmentation annotations differ far more between
datasets than detection ones do, so each dataset passes on what it reads:
polygons in `$segmentation` for COCO, `$polygon`, `$color`, `$semantic` and
`$instance` for Cityscapes depending on the requested `target_type`, one-hot
`$masks` for Oxford-IIIT Pet and Pascal VOC. `$labels` is not uniform either:
the detection datasets name their classes, while
`pascal_segmentation_dataset()` numbers the planes of `$masks` that are not
empty.

What `draw_segmentation_masks()` and the segmentation models want is `$masks`,
an `(N, H, W)` tensor with one plane per annotated class. Two target transforms
build it from raw annotations:

| transform | reads |
|---|---|
| `target_transform_coco_masks()` | `$segmentation`, the COCO polygons |
| `target_transform_trimap_masks()` | `$trimap`, the three-value Oxford-IIIT Pet mask |

```{r}
seg_ds <- coco_segmentation_dataset(
  train = FALSE,
  download = TRUE,
  transform = transform_to_tensor,
  target_transform = target_transform_coco_masks
)
masked <- draw_segmentation_masks(seg_ds[1])
```

`coco_segmentation_dataset()` only promotes its items to
`image_with_segmentation_mask` once `$masks` is there, so without that target
transform they stay `image_with_bounding_box`. `cityscapes_dataset()` has no
target transform of its own, and its targets have to be turned into `$masks` by
hand before they can be drawn.

## Which function takes which class

| family | dispatches on | acts on |
|---|---|---|
| `transform_*()` | `torch_tensor`, `magick-image`, `array`, `matrix` | the image alone |
| `target_transform_*()` | `object_detection_target`, `segmentation_target` | the target alone |
| `item_transform_*()` | `image_with_bounding_box`, `image_with_rotated_box`, `image_with_segmentation_mask`, `dataset` | image and target together |
| `draw_bounding_boxes()` | `torch_tensor`, `image_with_bounding_box`, `image_with_rotated_box` | draws boxes on the image |
| `draw_segmentation_masks()` | `torch_tensor`, `image_with_segmentation_mask` | overlays masks on the image |

`draw_keypoints()` is not generic: it takes an image tensor and an `(N, K, 2)`
keypoint tensor.

Calling any of these on something they have no method for stops with an error
naming the class it was given, which is usually a sign that a target lost its
class somewhere along a pipeline.

## Building an item by hand

A bare list is not accepted as a target, so annotations that do not come out of
a torchvision dataset need their classes set before any of this will look at
them:

```{r}
img <- torch_randint(0, 255, size = c(3, 120, 160))$to(dtype = torch_uint8())

y <- list(
  boxes = torch_tensor(rbind(c(10, 20, 60, 90), c(70, 30, 150, 110)), dtype = torch_float()),
  labels = c("cat", "dog"),
  image_height = 120,
  image_width = 160
)
class(y) <- c("object_detection_target", "list")

item <- list(x = img, y = y)
class(item) <- c("image_with_bounding_box", "list")

str(item)
#> List of 2
#>  $ x:Byte [1:3, 1:120, 1:160]
#>  $ y:List of 4
#>   ..$ boxes       :Float [1:2, 1:4]
#>   ..$ labels      : chr [1:2] "cat" "dog"
#>   ..$ image_height: num 120
#>   ..$ image_width : num 160
#>   ..- attr(*, "class")= chr [1:2] "object_detection_target" "list"
#>  - attr(*, "class")= chr [1:2] "image_with_bounding_box" "list"
```

A segmentation target is built the same way, with `segmentation_target` and
`image_with_segmentation_mask`.

## What models return

Model outputs are plain lists and carry no class of their own. The object
detection models return `$detections`, one element per image of the batch, each
with `$boxes`, `$labels` and `$scores`; the semantic segmentation models return
`$out`, the per-class logits. The "Visualization utilities" article covers
turning either of them into something the drawing helpers can render.
